Single-fluorophore orientation determination with multiview polarized illumination: modeling and microscope design

### 2017-08-13 paper/paper/paper.tex
Hari's edits.

%%%%%%%%%%%%%%%%%%% abstract and OCIS codes %%%%%%%%%%%%%%%%
\begin{abstract*}
  We investigate the use of polarized illumination in multiview microscopes for
  determining the orientation of single fluorescent molecules. First, we relate
  the orientation of single molecules to measurable intensities in multiview
  microscopes and develop an information-theoretic metric---the solid-angle
  uncertainty---to compare the ability of multiview microscopes to estimate the
  orientation of single molecules. Next, we compare a broad class of microscopes
  using this metric---single- and dual-view microscopes with varying
  illumination polarization, illumination numerical aperture (NA), detection NA,
  obliquity, asymmetry, and exposure. We find that multiview microscopes are
  better at determining the orientation of single molecules than single-view
  microscopes. We also find that choosing a small illumination NA and a large
  detection NA are good design choices, that multiview widefield microscopes can
  benefit from oblique illumination and detection, and that asymmetric NA
  microscopes can benefit from exposure asymmetry.
\end{abstract*}

@@

\textbf{Spatial variation of illumination polarization} techniques vary the
polarization and intensity in the illumination path to excite molecules in
specific orientations in the focal volume \cite{debarre2004}. Although these
techniques can accurately estimate the orientation of molecules in all
orientations, they require scanning which makes these techniques too slow for
live cell imaging.

@@

To develop a polarized illumination or detection technique we need to choose a
method for varying the polarization and place it in the illumination or
detection path. The simplest and most inexpensive way to vary the polarization
is with a division-of-time (DoT) polarizer using a rotating polarizer or a
universal compensator/polarizer. Other polarization methods
(divison-of-amplitude, division-of-focal-plane) are more expensive and require
additional alignment or interpolation \cite{tyo2006}. The simplicity of DoT
polarization comes at the expense of temporal resolution---intensity
measurements are made one after the other which increases the acquisition
time. If we place the DoT polarizer in the detection path, we will block
valuable photons from reaching the detector which unnecessarily increases
exposure to the sample \cite{demay2011}. By placing the DoT polarizer in the
illumination arm we minimize exposure to the sample. We focus on DoT polarized
illumination in this paper because of its combination of simplicity and low
sample exposure.

@@

Recently, there has been increasing interest in multiview microscopy techniques
for biological imaging \cite{wu2013, keller2015, wu2016, wu2017}. Multiview
microscopes offer two major advantages over single-view microscopes. First,
multiview microscopes can achieve nearly isotropic resolution compared to the
anisotropic resolution of single-view microscopes whose axial resolution is
reduced by at least a factor of two compared to lateral resolution. Second,
multiview microscopes can use light-sheet illumination to reduce phototoxicity
without requiring dedicated illumination optics. For example, if two orthogonal
objectives are focused on the same point, then the objectives can alternate
roles as the light-sheet illumination path and as the detection
path. Light-sheet illumination offers a major reduction of phototoxicity because
a light sheet only illuminates in-focus regions while other widefield and
confocal approaches inevitably illuminate out-of-focus regions. Together, these
advantages make multiview microscopes good candidates for imaging live
biological specimens for long periods.

@@

We consider a single molecule with an absorption dipole moment
$\hat{\bs{\mu}}_{\text{abs}}$ that we express in spherical coordinates $\Theta$
and $\Phi$ (see Figure \ref{fig:coordinates}a)) as
\begin{align}
  \hat{\bs{\mu}}_{\text{abs}}(\Theta, \Phi) = \sin\Theta\cos\Phi\hat{\mb{x}} + \sin\Theta\sin\Phi\hat{\mb{y}} + \cos\Theta\hat{\mb{z}} \label{eq:mu}.
\end{align}
We place the molecule in the focal plane of an ideal, aplanatic,
polarization-preserving objective lens with its optical axis aligned with the
$\mh{z}$ axis. Next, we place a polarizer in the aperture plane with a variable
transmission axis $\hat{\mb{p}}_{\text{exc}}$ that we express as
\begin{align}
 \hat{\mb{p}}_{\text{exc}}(\phi_{\text{exc}}) = \cos\phi_{\text{exc}}\hat{\mb{x}} +
\sin\phi_{\text{exc}}\hat{\mb{y}},\label{eq:excfield}
\end{align}
where $\phi_{\text{exc}}$ is the angle between the transmission axis of the
polarizer and the positive $\mh{x}$ axis (see Figure
\ref{fig:coordinates}b)). Finally, we place a spatially incoherent light source
(or its image) in the aperture plane to establish widefield illumination. In
this geometry each point in the sample is illuminated by a range of angles that
can be changed by adjusting the radius of the stop in the aperture plane. We
denote the set of illuminated angles by $\Omega$ (see Figure
\ref{fig:coordinates}d)). If we use unit vectors $\hat{\mb{r}}$ expressed in
spherical coordinates $\theta$ and $\phi$ (see Figure \ref{fig:coordinates}c))
as
  \begin{align}
   \hat{\mb{r}}(\theta, \phi) = \sin\theta\cos\phi\hat{\mb{x}} + \sin\theta\sin\phi\hat{\mb{y}} + \cos\theta\hat{\mb{z}}, 
  \end{align}
  then the set of angles $\Omega$ can be expressed as
  \begin{align}
  \Omega = \{\phi, \theta\ |\ 0 < \phi \leq 2\pi, 0 < \theta \leq \alpha\},
  \end{align}
  where $\alpha$ is the angle between the optical axis and the most oblique
  illuminating ray. Equivalently, $\alpha$ can be expressed in terms of NA using
  \begin{align}
    \text{NA} = n\sin\alpha, 
  \end{align}
  where $n$ is the index of refraction of the sample medium. Note that the NA of
  the objective is always greater than or equal to the illumination NA because
  we can underfill the aperture plane.

@@

    To find the absorption efficiency of a single molecule under widefield
    illumination, we integrate over all illumination rays and divide by the
    total incident power which gives the vector expression
\begin{align}
  \eta_{\text{abs}} = \frac{\int_{\Omega}d\mh{r}|\hat{\bs{\mu}}_{\text{abs}}\cdot\mb{R}(\mh{r})\mh{p}_{\text{exc}}|^2}{\int_{\Omega}d\mh{r}}\label{eq:abs}. 
\end{align}

@@

We can extend these results to light-sheet illumination created by scanning a
laser. If we shine a narrow collimated laser beam onto a galvanometer in the
aperture plane and scan the galvanometer, we create a weakly focused light sheet
in the sample. If we underfill the aperture plane and scan the beam in the
sample with a velocity much less than the beam waist radius divided by the
coherence time, we can ignore longitudinal excitation and molecules in the
sample will be excited as if a single plane wave was incident. Therefore, we can
find the absorption efficiency of a molecule under weakly focused scanned laser
illumination by taking the limit of equation \ref{eq:scalarabs} as
$\alpha \rightarrow 0$ (or by plugging equation \ref{eq:mu} and
\ref{eq:excfield} into equation \ref{eq:singleray}) giving
\begin{align}
  \eta_{\text{abs}}^{\text{ray}} &= \sin^2\Theta\cos^2(\Phi - \phi_{\text{exc}}),
\end{align}
which is recognizable as Malus' law generalized to three dimensions. This means
that light-sheet illumination is approximately equivalent to low-NA widefield
illumination with respect to the molecular orientations they excite. Notice that
the absorption efficiency can take its maximum range of values
($\eta_{\text{abs}} \in [0, 1]$) when $\alpha=0$.

@@

\subsection{Intensity Measurements}\label{forward}
The detected intensity is the product of the the sample exposure
$I_{\text{tot}}$, the absorption efficiency, and the detection efficiency
\begin{align}
  I = I_{\text{tot}}\eta_{\text{abs}}\eta_{\text{det}}\label{eq:forward_frame}.
\end{align}
Equation \ref{eq:forward_frame} is the forward model for a single intensity
measurement. Figure \ref{fig:single-frame} shows the efficiencies in three
representative geometries.

A single intensity measurement does not give us enough information to
reconstruct the orientation of a single molecule. Next, we will consider
combining several intensity measurements to create complete polarized
illumination microscopes.

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{single-frame}
\caption{Representative examples of single intensity measurements.\newline
  \newline \textbf{Columns left to right:} 1) schematics where the solid line
  encloses the illumination solid angle, the dashed line encloses the detection
  solid angle, and the arrow indicates the transmission axis of the illumination
  polarizer; 2) the absorption efficiency, see equation \ref{eq:scalarabs}; 3)
  the detection efficiency, see equation \ref{eq:scalardet}; (4) the total
  efficiency, the product of the absorption and detection efficiencies.\newline
  \newline \textbf{Rows top to bottom:} 1) coincident illumination
  ($\text{NA} = 1.1$ with $x$-polarized light) and detection (NA = 1.1); 2)
  non-coincident orthogonal illumination ($\text{NA} = 0.8$) and detection
  ($\text{NA} = 0.8$); 3) non-coincident 135${}^{\circ}$-separated illumination
  ($\text{NA} = 0.8$) and detection ($\text{NA} = 0.8$). All simulations use
  $n=1.33$.}
  \label{fig:single-frame}
\end{figure}

@@

Figures \ref{fig:symmetric-widefield} and \ref{fig:double-arm} show our results
for dual-view symmetric widefield designs. We illuminate from one objective and
detect from the other for four polarization orientations, then we repeat the
polarization orientations with the same objectives in reversed roles. In Figure
\ref{fig:symmetric-widefield} we sweep through the NA of both objectives and the
angle between the objectives while considering steric constraints. Note how
adding a second view removes the high uncertainty region in the transverse
plane. The dual-view microscope still has high uncertainty regions near the
optical axes, but the uncertainty is reduced by almost two orders of magnitude
compared to the single-view microscope. We find that the lowest median of the
solid-angle uncertainty occurs with the largest possible NA objectives. We also
find that increasing NA always lowers the MAD, but orthogonal arms are not
always best. At large NA, it is advantageous to move the objectives together or
against the cover slip. Given the high uncertainty in the transverse plane for a
single-view shown in Figure \ref{fig:single-arm}b), we can see why oblique views
perform better than orthogonal views---oblique views are complementary because
the regions of high orientation uncertainty from each view do not overlap.

\begin{figure}[htbp]
\centering\includegraphics[width=\textwidth]{symmetric-widefield}
\caption{Dual-view symmetric widefield designs with varying NA and angle between
  objectives. a) Schematic of the microscope. We illuminate with the first objective
  (red solid) and detect from the second objective (red dashed). Then we
  illuminate from the second objective (blue solid) and detect from the first
  objective (blue dashed). b) Solid-angle uncertainty for the microscope in
  a). c) Median of the solid-angle uncertainty as a function of NA and the angle
  between the objectives. d) MAD of the solid-angle uncertainty as a function of NA
  and the angle between the objectives. The microscope in a) and b) is indicated
  by a cross in c) and d). e) Median of the solid-angle uncertainty as a
  function of the angle between the objectives when NA=0.6. f) MAD of the
  solid-angle uncertainty as a function of the angle between the objectives when
  NA=0.6. The profile in e) and f) is taken along the green line in c) and d).}
\label{fig:symmetric-widefield}
\end{figure}

Figure \ref{fig:double-arm} shows our results when we used a dual-view symmetric
orthogonal widefield design and varied the illumination and detection NA. Our
results are similar to the single-view case in Figure \ref{fig:single-arm}. We
find that the best designs use a small illumination NA and a large detection NA
where $\alpha \approx 57.4^{\circ}$. As we discussed in Section
\ref{excitation}, low-NA widefield illumination is approximately equivalent to
light-sheet illumination for the task of estimating the orientation of
molecules. Therefore, the preference for low-NA widefield illumination also
implies that dual-view light-sheet illumination geometries are an excellent
choice for uniformly reconstructing the orientation of single molecules.

\begin{figure}[htbp]
\centering\includegraphics[width=\textwidth]{double-arm}
\caption{Dual-view symmetric orthogonal widefield designs with varying
  illumination and detection NA. a) Schematic of the microscope. b) Solid-angle
  uncertainty for the microscope in a). c) Median of the solid-angle uncertainty
  as a function of illumination and detection NA. d) MAD of the solid-angle
  uncertainty as a function of illumination and detection NA. The microscope in
  a) and b) is indicated by a cross in c) and d).}
\label{fig:double-arm}
\end{figure}

Figure \ref{fig:asymmetric-double} shows our results for dual-view asymmetric
light-sheet illumination designs. We used light-sheet illumination on both sides
and kept the objectives orthogonal so that the light sheet from one objective
illuminates the focal plane of the other objective. Then, we swept through the
NA of one objective while maximizing the NA of the other objective so that both
objectives would touch the cover slip and each other
($\alpha_1 + \alpha_2 = 90^{\circ}$ in Figure \ref{fig:asymmetric-double}a)). We
swept through the NA asymmetry and the sample-exposure asymmetry while keeping
the total sample exposure constant, and found that symmetric designs are at a
local minimum of the median and MAD of the solid-angle uncertainty. We also
found that at extreme asymmetries where the exposure from the low NA arm is much
larger than the exposure from the high NA arm (top-left and bottom-right corners
of Figures \ref{fig:asymmetric-double}c) and d)) the median and MAD are
comparable to the symmetric light-sheet microscope. This means that if a very
high NA objective is available and a low NA objective can provide light sheet
illumination, this microscope will perform comparably with a dual-view
microscope with two objectives of medium NA. Figure \ref{fig:asymmetric-double}
also shows that designs with slightly asymmetric NA can trade off a low
solid-angle uncertainty median for a low solid-angle uncertainty MAD by changing
the sample exposure ratio.

@@

Table 1 shows a summary of our evaluation metrics for microscopes in each
class. The single-view widefield microscope with a large illumination NA and
high detection NA has the lowest solid-angle uncertainty median but the largest
max. The dual-view oblique symmetric widefield with an intermediate NA performs
reasonably well on all three metrics, while the dual-view orthogonal symmetric
light-sheet microscopes with large NA perform very well on all three metrics.

\begin{table}[ht!]
\centering
\caption{Comparison of designs in each class of microscope. All values are in
  steradians.}
\begin{tabular}{llll}
  \toprule
Microscope Type&Max\{$\sigma_{\Omega}$\}&Median\{$\sigma_{\Omega}$\}&MAD\{$\sigma_{\Omega}$\}\\ \midrule
Single-view widefield (NA${}_\textrm{ill}$=0, NA${}_\textrm{det}$=1.1)&3.04$\times 10^{0}$&8.20$\times 10^{-4}$&1.48$\times 10^{-4}$ \\ 
  Dual-view oblique symmetric widefield (NA=0.6, $\beta$=53${}^{\circ}$)&1.35$\times 10^{-1}$&4.49$\times 10^{-3}$&9.17$\times 10^{-4}$\\
Dual-view orthogonal symmetric light-sheet (NA=0.8)&6.23$\times 10^{-3}$&1.79$\times 10^{-3}$&1.65$\times 10^{-4}$\\  
Dual-view orthogonal symmetric light-sheet (NA=0.94)&3.13$\times 10^{-3}$&1.28$\times 10^{-3}$&1.14$\times 10^{-4}$\\
\bottomrule
\end{tabular}
\end{table}

\section{Discussion and Conclusions}\label{discussion}
In this work we have developed a model to predict the intensities measured by
polarized illumination microscopes in a wide range of geometries, developed
metrics to measure the ability of a microscope design to estimate the
orientation of single molecules, and used these metrics to compare a wide
range of microscope designs. Our main result is a short list of design
heuristics that can be used to design polarized-illumination single-molecule
orientation microscopes:
\begin{itemize}
\item All microscopes benefit from a small illumination NA and a large detection
  NA with $\alpha \approx 57.4^{\circ}$.
\item Dual-view microscopes outperform single-view microscopes, and dual-view
  microscopes remove the high uncertainty region in the transverse plane of
  single-view microscopes.
\item High NA dual-view widefield microscopes are best when the two views are
  oblique. Oblique views are complementary because their high uncertainty
  transverse planes do not overlap.
\item Symmetric NA dual-view orthogonal light-sheet microscopes provide a good
  compromise between cost, solid-angle uncertainty median, and solid-angle
  uncertainty MAD. Asymmetric NA dual-view orthogonal light-sheet microscopes
  can improve the solid-angle uncertainty median or MAD by changing the sample
  exposure ratio between the two views.
\end{itemize}

Our results are limited in several ways. First, we only consider single
molecules in homogeneous environments. We ignore light reflected from the cover
slip and any other inhomogeneities in the sample. If we considered light
reflected from the cover slip we expect that the results would put a lower
weighting on large detection NAs because we would collect more light. Second, we
only considered ideal, aplanatic, polarization-preserving objectives. In
practice, these conditions are not always satisfied and the power from high NA
rays is apodized by a factor of $\frac{n}{\cos\Theta}$
\cite{nov2006}. Therefore, our results are most accurate for low NA rays and get
progressively worse for high NA rays, although our major conclusions would not
change if we included this apodization factor. The main advantage of ignoring
apodization is the tractability of the problem---we have provided approximate
closed-form solutions for the illumination and detection efficiencies that have
allowed us to draw useful design conclusions. Third, we have used a Poisson
noise model which ignores noise introduced by the detector. Detector noise can
be a large fraction of the total noise at low intensities, so including detector
noise would increase the informational value of high intensity
measurements. Therefore, we expect that including the effects of detector noise
would favor microscopes with a large detection NA. Finally, we have focused on
the task of estimating the orientation of molecules and ignored the equally
important task of estimating their location. The task of jointly estimating the
orientation and location of single molecules from multiple views is a topic for
future work.

@@

\section*{Funding}
TODO

>>>>

%%%%%%%%%%%%%%%%%%% abstract and OCIS codes %%%%%%%%%%%%%%%%
\begin{abstract*}
  We investigate the use of polarized illumination in multiview microscopes for
  determining the orientation of single fluorescent molecules. First, we relate
  the orientation of single molecules to measurable intensities in multiview
  microscopes and develop an information-theoretic metric---the solid-angle
  uncertainty---to compare the ability of multiview microscopes to estimate the
  orientation of single molecules. Next, we compare a broad class of microscopes
  using this metric---single- and dual-view microscopes with varying
  illumination polarization, illumination numerical aperture (NA), detection NA,
  obliquity, asymmetry, and exposure. We find that multiview microscopes are
  better at determining the orientation of single molecules than single-view
  microscopes. We also find that choosing a small illumination NA and a large
  detection NA are good design choices, that multiview K\"ohler illumination
  microscopes can benefit from oblique illumination and detection, and that
  asymmetric NA microscopes can benefit from exposure asymmetry.
\end{abstract*}

@@

\textbf{Spatial variation of illumination polarization} techniques vary the
polarization and intensity in the illumination path to excite molecules in
specific orientations in the focal volume \cite{debarre2004}. Although these
techniques can accurately estimate the orientation of molecules in all
orientations, they require scanning which makes acquisition time a concern in
live-cell imaging.

@@

To develop a polarized illumination or detection technique we need to choose a
method for varying the polarization and place it in the illumination or
detection path. The simplest and most inexpensive way to vary the polarization
is with a division-of-time (DoT) polarizer using a rotating polarizer or a
universal compensator/polarizer. Other polarization methods
(divison-of-amplitude, division-of-focal-plane) are more expensive and require
additional alignment or interpolation \cite{tyo2006}. The simplicity of DoT
polarization comes at the expense of temporal resolution---intensity
measurements are made one after the other which increases the acquisition
time. If we place the DoT polarizer in the detection path, we will block
valuable photons from reaching the detector which leads to unnecessary exposure
to the sample \cite{demay2011}. By placing the DoT polarizer in the illumination
arm we minimize exposure to the sample. We focus on DoT polarized illumination
in this paper because of its combination of simplicity and low sample exposure.

@@

Recently, there has been increasing interest in multiview microscopy techniques
for biological imaging \cite{wu2013, keller2015, wu2016, wu2017}. Multiview
microscopes offer two major advantages over single-view microscopes. First,
multiview microscopes can achieve nearly isotropic resolution compared to the
anisotropic resolution of single-view microscopes whose axial resolution is
reduced by at least a factor of two compared to lateral resolution. Second,
multiview microscopes can use light-sheet illumination to reduce phototoxicity
without requiring a dedicated illumination path. For example, if two orthogonal
objectives are focused on the same point, then the objectives can alternate
roles as the light-sheet illumination path and as the detection
path. Light-sheet illumination offers a major reduction of phototoxicity because
a light sheet only illuminates in-focus regions while other widefield and
confocal approaches inevitably illuminate out-of-focus regions. Together, these
advantages make multiview microscopes good candidates for imaging live
biological specimens for long periods.

@@

We consider a single molecule with an absorption dipole moment
$\hat{\bs{\mu}}_{\text{abs}}$ that we express in spherical coordinates $\Theta$
and $\Phi$ (see Figure \ref{fig:coordinates}a)) as
\begin{align}
  \hat{\bs{\mu}}_{\text{abs}}(\Theta, \Phi) = \sin\Theta\cos\Phi\hat{\mb{x}} + \sin\Theta\sin\Phi\hat{\mb{y}} + \cos\Theta\hat{\mb{z}} \label{eq:mu}.
\end{align}
We place the molecule in the focal plane of an ideal, aplanatic,
polarization-preserving objective lens with its optical axis aligned with the
$\mh{z}$ axis. Next, we place a polarizer in the aperture plane with a variable
transmission axis $\hat{\mb{p}}_{\text{exc}}$ that we express as
\begin{align}
 \hat{\mb{p}}_{\text{exc}}(\phi_{\text{exc}}) = \cos\phi_{\text{exc}}\hat{\mb{x}} +
\sin\phi_{\text{exc}}\hat{\mb{y}},\label{eq:excfield}
\end{align}
where $\phi_{\text{exc}}$ is the angle between the transmission axis of the
polarizer and the positive $\mh{x}$ axis (see Figure
\ref{fig:coordinates}b)). Finally, we place a spatially incoherent and spatially
uniform light source (or its image) in the aperture plane to establish K\"ohler
illumination. In this geometry each point in the sample is illuminated by a
range of angles that can be changed by adjusting the radius of the stop in the
aperture plane. We denote the set of illuminated angles by $\Omega$ (see Figure
\ref{fig:coordinates}d)). If we use unit vectors $\hat{\mb{r}}$ expressed in
spherical coordinates $\theta$ and $\phi$ (see Figure \ref{fig:coordinates}c))
as
  \begin{align}
   \hat{\mb{r}}(\theta, \phi) = \sin\theta\cos\phi\hat{\mb{x}} + \sin\theta\sin\phi\hat{\mb{y}} + \cos\theta\hat{\mb{z}}, 
  \end{align}
  then the set of angles $\Omega$ can be expressed as
  \begin{align}
  \Omega = \{\phi, \theta\ |\ 0 < \phi \leq 2\pi, 0 < \theta \leq \alpha\},
  \end{align}
  where $\alpha$ is the angle between the optical axis and the most oblique
  illuminating ray. Equivalently, $\alpha$ can be expressed in terms of NA using
  \begin{align}
    \text{NA} = n\sin\alpha, 
  \end{align}
  where $n$ is the index of refraction of the sample medium. Note that the NA of
  the objective is always greater than or equal to the illumination NA because
  we can underfill the aperture plane.

@@

    To find the absorption efficiency of a single molecule under K\"ohler
    illumination, we integrate over all illumination rays and divide by the
    total incident power which gives the vector expression
\begin{align}
  \eta_{\text{abs}} = \frac{\int_{\Omega}d\mh{r}|\hat{\bs{\mu}}_{\text{abs}}\cdot\mb{R}(\mh{r})\mh{p}_{\text{exc}}|^2}{\int_{\Omega}d\mh{r}}\label{eq:abs}. 
\end{align}

@@

We can extend these results to light-sheet illumination created by scanning a
laser. If we shine a narrow collimated laser beam onto a galvanometer in the
aperture plane and scan the galvanometer, we create a weakly focused light sheet
in the sample. If we underfill the aperture plane and scan the beam in the
sample with a velocity much less than the beam waist radius divided by the
coherence time, we can ignore the excitation caused by fields aligned with the
optical axis. In this case, molecules in the sample will be excited as if a
single plane wave was incident. Therefore, we can find the absorption efficiency
of a molecule under weakly focused scanned laser illumination by taking the
limit of equation \ref{eq:scalarabs} as $\alpha \rightarrow 0$ (or by plugging
equation \ref{eq:mu} and \ref{eq:excfield} into equation \ref{eq:singleray})
giving
\begin{align}
  \eta_{\text{abs}}^{\text{ray}} &= \sin^2\Theta\cos^2(\Phi - \phi_{\text{exc}}),
\end{align}
which is recognizable as Malus' law generalized to three dimensions. This means
that light-sheet illumination is approximately equivalent to low-NA K\"ohler
illumination with respect to the molecular orientations they excite. Notice that
the absorption efficiency can take its maximum range of values
($\eta_{\text{abs}} \in [0, 1]$) when $\alpha=0$.

@@

\subsection{Intensity Measurements}\label{forward}
The detected intensity is the product of the the sample exposure
$I_{\text{tot}}$, the absorption efficiency, and the detection efficiency
\begin{align}
  I = I_{\text{tot}}\eta_{\text{abs}}\eta_{\text{det}}\label{eq:forward_frame}.
\end{align}
Equation \ref{eq:forward_frame} is the forward model for a single intensity
measurement. Figure \ref{fig:single-frame} shows the efficiencies in three
representative geometries. Note that in all cases there are multiple
orientations that give the same intensity measurement. This means that a single
intensity measurement does not give us enough information to reconstruct the
orientation of a single molecule. Next, we will consider combining several
intensity measurements to create complete polarized illumination microscopes.

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{single-frame}
\caption{Representative examples of single intensity measurements. Black dots
  indicate where the Cartesian unit vectors intersect the unit sphere. \newline \newline \textbf{Columns left to right:} 1) schematics where
  the solid line encloses the illumination solid angle, the dashed line encloses
  the detection solid angle, and the arrow indicates the transmission axis of
  the illumination polarizer; 2) the absorption efficiency, see equation
  \ref{eq:scalarabs}; 3) the detection efficiency, see equation
  \ref{eq:scalardet}; (4) the total efficiency, the product of the absorption
  and detection efficiencies.\newline \newline \textbf{Rows top to bottom:} 1)
  coincident illumination ($\text{NA} = 1.1$ with $x$-polarized light) and
  detection (NA = 1.1); 2) non-coincident orthogonal illumination
  ($\text{NA} = 0.8$) and detection ($\text{NA} = 0.8$); 3) non-coincident
  135${}^{\circ}$-separated illumination ($\text{NA} = 0.8$) and detection
  ($\text{NA} = 0.8$). All simulations use $n=1.33$.}
  \label{fig:single-frame}
\end{figure}

@@

Figures \ref{fig:symmetric-widefield} and \ref{fig:double-arm} show our results
for dual-view symmetric K\"ohler designs. We illuminate from one objective and
detect from the other for four polarization orientations, then we repeat the
polarization orientations with the same objectives in reversed roles. In Figure
\ref{fig:symmetric-widefield} we sweep through the NA of both objectives and the
angle between the objectives while considering steric constraints. Note how
adding a second view removes the high uncertainty region in the transverse
plane. The dual-view microscope still has high uncertainty regions near the
optical axes, but the uncertainty is reduced by almost two orders of magnitude
compared to the single-view microscope. We find that the lowest median of the
solid-angle uncertainty occurs with the largest possible NA objectives. We also
find that increasing NA always lowers the MAD, but orthogonal arms are not
always best. At large NA, it is advantageous to move the objectives together or
against the cover slip. Given the high uncertainty in the transverse plane for a
single-view shown in Figure \ref{fig:single-arm}b), we can see why oblique views
perform better than orthogonal views---oblique views are complementary because
the regions of high orientation uncertainty from each view do not overlap.

\begin{figure}[htbp]
\centering\includegraphics[width=\textwidth]{symmetric-widefield}
\caption{Dual-view symmetric K\"ohler designs with varying NA and angle between
  objectives. a) Schematic of the microscope. We illuminate with the first objective
  (red solid) and detect from the second objective (red dashed). Then we
  illuminate from the second objective (blue solid) and detect from the first
  objective (blue dashed). b) Solid-angle uncertainty for the microscope in
  a). c) Median of the solid-angle uncertainty as a function of NA and the angle
  between the objectives. d) MAD of the solid-angle uncertainty as a function of NA
  and the angle between the objectives. The microscope in a) and b) is indicated
  by a cross in c) and d). e) Median of the solid-angle uncertainty as a
  function of the angle between the objectives when NA=0.6. f) MAD of the
  solid-angle uncertainty as a function of the angle between the objectives when
  NA=0.6. The profile in e) and f) is taken along the orange line in c) and d).}
\label{fig:symmetric-widefield}
\end{figure}

Figure \ref{fig:double-arm} shows our results when we used a dual-view symmetric
orthogonal K\"ohler design and varied the illumination and detection NA. Our
results are similar to the single-view case in Figure \ref{fig:single-arm}. We
find that the best designs use a small illumination NA and a large detection NA
where $\alpha \approx 57.4^{\circ}$. As we discussed in Section
\ref{excitation}, low-NA K\"ohler illumination is approximately equivalent to
light-sheet illumination for the task of estimating the orientation of
molecules. Therefore, the preference for low-NA K\"ohler illumination also
implies that dual-view light-sheet illumination geometries are an excellent
choice for uniformly reconstructing the orientation of single molecules.

\begin{figure}[htbp]
\centering\includegraphics[width=\textwidth]{double-arm}
\caption{Dual-view symmetric orthogonal K\"ohler designs with varying
  illumination and detection NA. a) Schematic of the microscope. b) Solid-angle
  uncertainty for the microscope in a). c) Median of the solid-angle uncertainty
  as a function of illumination and detection NA. d) MAD of the solid-angle
  uncertainty as a function of illumination and detection NA. The microscope in
  a) and b) is indicated by a cross in c) and d).}
\label{fig:double-arm}
\end{figure}

Figure \ref{fig:asymmetric-double} shows our results for dual-view asymmetric
light-sheet illumination designs. We used light-sheet illumination on both sides
and kept the objectives orthogonal so that the light sheet from one objective
illuminates the focal plane of the other objective. Then, we swept through the
NA of one objective while maximizing the NA of the other objective so that both
objectives would touch the cover slip and each other
($\alpha_1 + \alpha_2 = 90^{\circ}$ in Figure \ref{fig:asymmetric-double}a)). We
swept through the NA asymmetry and the sample-exposure asymmetry while keeping
the total sample exposure constant, and found that symmetric designs are at a
local minimum of the median and MAD of the solid-angle uncertainty. We also
found that at extreme asymmetries where the exposure from the low-NA arm is much
larger than the exposure from the high-NA arm (top-left and bottom-right corners
of Figures \ref{fig:asymmetric-double}c) and d)) the median and MAD are
comparable to the symmetric light-sheet microscope. This means that if a very
high-NA objective is available and a low-NA objective can provide light sheet
illumination, this microscope will perform comparably with a dual-view
microscope with two medium-NA objectives. Note that this asymmetric exposure
configuration maximizes the use of the high-NA detection objective---if a
high-NA objective is available it should be used for detection as much as
possible. Figure \ref{fig:asymmetric-double} also shows that designs with
slightly asymmetric NA can trade off a low solid-angle uncertainty median for a
low solid-angle uncertainty MAD by changing the sample exposure ratio.

@@

Table 1 shows a summary of our evaluation metrics for microscopes in each
class. The single-view K\"ohler microscope with a small illumination NA and
high detection NA has the lowest solid-angle uncertainty median but the largest
max. The dual-view oblique symmetric K\"ohler with an intermediate NA performs
reasonably well on all three metrics, while the dual-view orthogonal symmetric
light-sheet microscopes with large NA perform very well on all three metrics.

\begin{table}[ht!]
\centering
\caption{Comparison of designs in each class of microscope. All values are in
  steradians.}
\begin{tabular}{llll}
  \toprule
Microscope Type&Max\{$\sigma_{\Omega}$\}&Median\{$\sigma_{\Omega}$\}&MAD\{$\sigma_{\Omega}$\}\\ \midrule
Single-view K\"ohler (NA${}_\textrm{ill}$=0, NA${}_\textrm{det}$=1.1)&3.04$\times 10^{0}$&8.20$\times 10^{-4}$&1.48$\times 10^{-4}$ \\ 
  Dual-view oblique symmetric K\"ohler (NA=0.6, $\beta$=53${}^{\circ}$)&1.35$\times 10^{-1}$&4.49$\times 10^{-3}$&9.17$\times 10^{-4}$\\
Dual-view orthogonal symmetric light-sheet (NA=0.8)&6.23$\times 10^{-3}$&1.79$\times 10^{-3}$&1.65$\times 10^{-4}$\\  
Dual-view orthogonal symmetric light-sheet (NA=0.94)&3.13$\times 10^{-3}$&1.28$\times 10^{-3}$&1.14$\times 10^{-4}$\\
\bottomrule
\end{tabular}
\end{table}

\section{Discussion and Conclusions}\label{discussion}
In this work we have developed a model to predict the intensities measured by
polarized illumination microscopes in a wide range of geometries, developed
metrics to measure the ability of a microscope design to estimate the
orientation of single molecules, and used these metrics to compare a wide
range of microscope designs. Our main result is a short list of design
heuristics that can be used to design polarized-illumination single-molecule
orientation microscopes:
\begin{itemize}
\item All microscopes benefit from a small illumination NA and a large detection
  NA with $\alpha \approx 57.4^{\circ}$.
\item Dual-view microscopes outperform single-view microscopes, and dual-view
  microscopes remove the high uncertainty region in the transverse plane of
  single-view microscopes.
\item High-NA dual-view K\"ohler microscopes are best when the two views are
  oblique. Oblique views are complementary because their high uncertainty
  transverse planes do not overlap.
\item Symmetric NA dual-view orthogonal light-sheet microscopes provide a good
  compromise between cost, solid-angle uncertainty median, and solid-angle
  uncertainty MAD. Asymmetric NA dual-view orthogonal light-sheet microscopes
  can improve the solid-angle uncertainty median or MAD by changing the sample
  exposure ratio between the two views.
\end{itemize}

Our results are limited in several ways. First, we only consider single
molecules in homogeneous environments. We ignore light reflected from the cover
slip and any other inhomogeneities in the sample. If we considered light
reflected from the cover slip we expect that the results would put a lower
weighting on large detection NAs because we would collect more light. Second, we
only considered ideal, aplanatic, polarization-preserving objectives. In
practice, these conditions are not always satisfied and the power from high-NA
rays is apodized by a factor of $\frac{n}{\cos\Theta}$
\cite{nov2006}. Therefore, our results are most accurate for low-NA rays and get
progressively worse for high-NA rays, although our major conclusions would not
change if we included this apodization factor. The main advantage of ignoring
apodization is the tractability of the problem---we have provided approximate
closed-form solutions for the illumination and detection efficiencies that have
allowed us to draw useful design conclusions. Third, we have used a Poisson
noise model which ignores noise introduced by the detector. Detector noise can
be a large fraction of the total noise at low intensities, so including detector
noise would increase the informational value of high intensity
measurements. Therefore, we expect that including the effects of detector noise
would favor microscopes with a large detection NA. Finally, we have focused on
the task of estimating the orientation of molecules and ignored the equally
important task of estimating their location. The task of jointly estimating the
orientation and location of single molecules from multiple views is a topic for
future work.

@@

\section*{Funding}
Support for this work was provided by the Intramural Research Programs of the
National Institute of Biomedical Imaging and Bioengineering.
